### 2024-03-27 tesis_Ashish/Chapter1/chapter1.tex
update 28 Mar

In particle physics, the luminosity of colliding beams is a crucial measurement as it provides information about the rate at which rare processes occur and enables precise measurements of particle properties such as cross-section and couplings to fermions and bosons. Luminosity depends on several variables, including the number of protons in each beam, the size and shape of the beams, and their relative positioning. A slight change in any of these variables can significantly alter the luminosity.
%which in turn affects the interpretation of collision data.
Therefore, it is crucial to measure the luminosity with high precision.
%The measurement of luminosity with the pixel detector at the CMS experiment is particularly important  as it is a highly precise detector that is capable of measuring luminosity with great accuracy. It provides information about the number of charged particles passing through the detector, which is directly proportional to the number of collisions. This information is then used to calculate the luminosity. 
Improving the uncertainty in luminosity measurements is essential for many physics analyses.
%including searches for new particles and measurements of the properties of known particles.
It is a dominant factor in the measurement of Higgs Boson cross sections and couplings to other gauge bosons and fermions projected for High luminosity (HL)-LHC era, as shown in Fig~\ref{fig:lum_unc}. As such, the uncertainty in cross-section and coupling measurements cannot be better than the uncertainty in luminosity measurement. Therefore, ongoing efforts are focused on improving the precision of luminosity measurements, with the pixel detector at CMS being a key component of this effort. By reducing the luminosity uncertainty, physicists can make more precise measurements of particle properties and potentially discover new physics beyond the Standard Model.

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.8\textwidth]{ashish_thesis/lumi_precision.png}
\caption[Phase II Uncertainty Projection]{ %
   Higgs boson production processes cross section projected uncertainty. In the most precisely measured Higgs boson production process, gluon fusion (ggH), luminosity uncertainty will dominate other experimental systematic uncertainty at the HL-LHC \cite{Dainese:2703572}.
}
\label{fig:lum_unc}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/lhc_schematic.png}
\caption[CERN Accelerator Complex]{%
    CERN accelerator complex consists of Linac 2, Proton Synchrotron Booster (PSB), Proton Synchrotron (PS), Super Proton Synchrotron (SPS) and Large Hadron Collider (LHC) which is the accelerator chain for proton physics at the LHC. LHC ring showing four interaction points where the ALICE, ATLAS, LHCb and CMS detectors are stationed \cite{lhc_complex}.
}
\label{fig:lhc}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.75\textwidth]{ashish_thesis/cms_exp.png}
\caption[CMS Experiment Design]{%
   CMS detector with all of the sub detectors showing silicon tracker, electromagnetic and hadron calorimeters and muon chambers \cite{CMS:2008xjf}. 
}
\label{fig:cms}
\end{figure}

@@

\begin{table}[h]
  \centering
  \caption[Integrated luminosity and its precision for Run1 and Run2]{Summary of center of mass energy, integrated luminosity, and systematic uncertainty per year in CMS.}
\begin{tabular}{c c c c}
  Year & Center of Mass Energy & Integrated Luminosity & Systematic Uncertainty \\
  & (TeV) & (fb$^{-1}$) & (\%) \\
\hline
2010 & 7 & 0.045 & 11 \\
2011 & 7 and 8 & 6.1 & 2.2 \\
2012 & 8 & 23.3 & 2.5 \\
2015 & 13 & 4.3 & 1.6 \\
2016 & 13 & 41.6 & 1.2 \\
2017 & 13 & 49.8 & 2.3 \\
2018 & 13 & 67.9 &  2.5\\
\end{tabular}
\label{tab:lumi}
\end{table}

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/CMS_luminosity.png}
\caption[CMS Integrated luminosity]{%
    Delivered luminosity versus time for Run 1, Run 2 and Run 3 of the CMS Experiment \cite{wikicern_cc}.
}
\label{fig:lumi}
\end{figure}

\begin{table}[h]
  \centering
  \caption[Precision for Run1 and Run2]{Summary of center of mass energy, integrated luminosity, and systematic uncertainty per year during the CMS pPb and PbPb luminosity measurement \cite{CMS-PAS-LUM-13-002, CMS-PAS-LUM-17-002, CMS-PAS-LUM-18-001}.}
\begin{tabular}{c c c c}
  Year & Center of Mass Energy/nucleon & Systematic Uncertainty \\
  & (TeV) & (\%) \\
\hline
2013 & 2.76 & 3.4(3.6)  \\
2016 &  5.02  & 3.2(3.7)\\
2018 & 5.02 & 1.5 \\
\end{tabular}
\label{tab:lumi40}
\end{table}

>>>>

In particle physics, the luminosity of colliding beams is a crucial measurement as it provides information about the rate at which rare processes occur and enables precise measurements of particle properties such as cross-section and couplings to fermions and bosons. Luminosity depends on several variables, including the number of protons in each beam, the size and shape of the beams, and their relative positioning. A slight change in any of these variables can significantly alter the luminosity.
%which in turn affects the interpretation of collision data.
Therefore, it is crucial to measure the luminosity with high precision.

Precision in luminosity measurement allows for the accurate determination of the Standard Model parameters, such as the masses of fundamental particles and the strength of their interactions (coupling constants). These parameters are critical for making precise predictions about particle behavior in high-energy collisions. By refining these parameters, physicists can enhance their understanding of the universe fundamental structure and forces.
High precision luminosity measurement increase the ability to observe rare processes that occur at very low rates. Such processes might be signatures of new physics beyond the Standard Model, including new particles or forces. Distinguishing these rare events from the background noise is crucial for advancing our understanding of the universe.
The electroweak theory, which unifies the weak nuclear force and electromagnetism, makes precise predictions about the production and decay rates of W and Z bosons. Accurate luminosity measurements are essential for comparing these theoretical predictions with experimental data. Any discrepancies could indicate new physics or the need for adjustments within the Standard Model.
Quantum Chromodynamics (QCD) describes the strong force, one of the four fundamental forces. Experiments that measure jet production rates and other phenomena related to the strong force depend on precise luminosity measurements for validating QCD predictions. This understanding is crucial for both the theoretical framework and for interpreting results of particle collision experiments.
High precision in luminosity measurement allows experimental results to be benchmarked against theoretical models with greater accuracy. This benchmarking is essential for validating or refuting theories about the fundamental constituents and forces of the universe.

%The measurement of luminosity with the pixel detector at the CMS experiment is particularly important  as it is a highly precise detector that is capable of measuring luminosity with great accuracy. It provides information about the number of charged particles passing through the detector, which is directly proportional to the number of collisions. This information is then used to calculate the luminosity. 
Improving the uncertainty in luminosity measurements is essential for many physics analyses.
%including searches for new particles and measurements of the properties of known particles.
It is a dominant factor in the measurement of Higgs Boson cross sections and couplings to other gauge bosons and fermions projected for High luminosity (HL)-LHC era, as shown in Fig~\ref{fig:lum_unc}. As such, the uncertainty in cross-section and coupling measurements cannot be better than the uncertainty in luminosity measurement. Therefore, ongoing efforts are focused on improving the precision of luminosity measurements, with the pixel detector at CMS being a key component of this effort. By reducing the luminosity uncertainty, physicists can make more precise measurements of particle properties and potentially discover new physics beyond the Standard Model.

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.8\textwidth]{ashish_thesis/lumi_precision.png}
\caption[Phase 2 uncertainty projection]{ %
   Higgs boson production processes cross section projected uncertainty. In the most precisely measured Higgs boson production process, gluon fusion (ggH), luminosity uncertainty will dominate other experimental systematic uncertainty at the HL-LHC \cite{Dainese:2703572}.
}
\label{fig:lum_unc}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/lhc_schematic.png}
\caption[CERN accelerator complex]{%
    CERN accelerator complex consists of Linac 2, Proton Synchrotron Booster (PSB), Proton Synchrotron (PS), Super Proton Synchrotron (SPS) and Large Hadron Collider (LHC) which is the accelerator chain for proton physics at the LHC. LHC ring showing four interaction points where the ALICE, ATLAS, LHCb and CMS detectors are stationed \cite{lhc_complex}.
}
\label{fig:lhc}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.75\textwidth]{ashish_thesis/cms_exp.png}
\caption[CMS experiment design]{%
   CMS detector with all of the sub detectors showing silicon tracker, electromagnetic and hadron calorimeters and muon chambers \cite{CMS:2008xjf}. 
}
\label{fig:cms}
\end{figure}

@@

\begin{table}[h]
  \centering
  \caption[Integrated luminosity and its precision per year]{Summary of center of mass energy, integrated luminosity, and systematic uncertainty per year in CMS.}
\begin{tabular}{c c c c}
  Year & Center of Mass Energy & Integrated Luminosity & Systematic Uncertainty \\
  & (TeV) & (fb$^{-1}$) & (\%) \\
\hline
2010 & 7 & 0.045 & 11 \\
2011 & 7 and 8 & 6.1 & 2.2 \\
2012 & 8 & 23.3 & 2.5 \\
2015 & 13 & 4.3 & 1.6 \\
2016 & 13 & 41.6 & 1.2 \\
2017 & 13 & 49.8 & 2.3 \\
2018 & 13 & 67.9 &  2.5\\
\end{tabular}
\label{tab:lumi}
\end{table}

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/CMS_luminosity.png}
\caption[CMS integrated luminosity]{%
    Delivered luminosity versus time for Run 1, Run 2 and Run 3 of the CMS Experiment \cite{wikicern_cc}.
}
\label{fig:lumi}
\end{figure}

\begin{table}[h]
  \centering
  \caption[Precision for pPb and PbPb collisions]{Summary of center of mass energy and systematic uncertainty per year during the CMS pPb and PbPb luminosity measurement \cite{CMS-PAS-LUM-13-002, CMS-PAS-LUM-17-002, CMS-PAS-LUM-18-001}.}
\begin{tabular}{c c c c}
  Year & Center of Mass Energy/nucleon & Systematic Uncertainty \\
  & (TeV) & (\%) \\
\hline
2013 & 2.76 & 3.4(3.6)  \\
2016 &  5.02  & 3.2(3.7)\\
2018 & 5.02 & 1.5 \\
\end{tabular}
\label{tab:lumi40}
\end{table}

### 2024-03-27 tesis_Ashish/Chapter2/chapter2.tex
update 28 Mar

where $L_{inst}$ is the luminosity, N is the number of events with zero particles detected in the region of interest, $N_0$ is the total number of events, $\sigma_{vis}$ is the visible cross-section. One advantage is that it is a simple and robust method that does not require sophisticated detectors or calibration procedures. However, it can be affected by background events, such as cosmic rays or beam-gas interactions. \\
The HFET algorithm measures luminosity by summing transverse energy deposits in the HF calorimeter cells. It relies on the direct proportionality between the total transverse energy and the particle flux from the interaction point. Accurate calibration of calorimeter cells is crucial for this method, as it ensures that the aggregate transverse energy accurately reflects the luminosity of the particle collisions \cite{Sirunyan:2759951.

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/luminometer_cms.png}
\caption[CMS Luminometers]{%
   CMS detector with all of the luminosity measurement devices showing PLT, BCM1F, HF, DT, PCC and RAMSES \cite{meyer2021luminosity}. 
}
\label{fig:lumino_cms}
\end{figure}

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/luminometer_cms_pseudo.png}
\caption[CMS luminometers Location]{%
    Cross section of the CMS detector in the r-z plane showing all Run 2 luminometers PLT, BCM1F, DTs, and HF location. The two RAMSES monitors used as a luminometer in Run 2 are located directly behind HF \cite{Sirunyan:2759951}. 
}
\label{fig:lumino_cms_pseudo}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/phaseI_upgrade_pixel_detector.png}
\caption[Pixel Detector Layout]{%
   Layout of the CMS Phase-1 upgraded pixel detector in longitudinal view \cite{Adam:2748381}. 
}
\label{fig:phaseI_upgrade}
\end{figure}

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/silicon_pixel_module.png}
\caption[Pixel Module Design]{%
   Pixel detector modules for BPIX L1 (left), BPIX L2-4 (middle), and
the FPIX detector (right) \cite{Adam:2748381}. 
}
\label{fig:pixelmodule}
\end{figure}

>>>>

where $L_{inst}$ is the luminosity, N is the number of events with zero particles detected in the region of interest, $N_0$ is the total number of events, $\sigma_{vis}$ is the visible cross-section. One advantage is that it is a simple and robust method that does not require sophisticated detectors or calibration procedures. However, it can be affected by background events, such as cosmic rays or beam-gas interactions. \\
The HFET algorithm measures luminosity by summing transverse energy deposits in the HF calorimeter cells. It relies on the direct proportionality between the total transverse energy and the particle flux from the interaction point. Accurate calibration of calorimeter cells is crucial for this method, as it ensures that the aggregate transverse energy accurately reflects the luminosity of the particle collisions \cite{Sirunyan:2759951}.

%The Hadron Forward Transverse Energy (HFET) algorithm is a method used to determine the luminosity of particle collisions. This method is based on the direct proportionality between the transverse energy deposited in the Hadron Forward calorimeter cells and the particle flux from the collision point.

The measurement of instantaneous luminosity, $L_{inst}$, using the HFET algorithm involves the following steps:

\begin{enumerate}
  \item Initialize the total transverse energy sum: 
  \[ \Sigma E_T = 0 \]

  \item For each calorimeter cell, do:
  \begin{enumerate}
    \item Measure the transverse energy \( E_{T,\text{cell}} \).
    \item Update the total transverse energy sum: 
    \[ \Sigma E_T = \Sigma E_T + E_{T,\text{cell}} \]
  \end{enumerate}

  \item Correct the total transverse energy sum for background energy:
  \[ \Sigma E_T = \Sigma E_T - E_{T,\text{background}} \]

  \item Multiply by a calibration constant \( k \) to obtain the luminosity:
  \[ L_{inst} = \( k \ \times \Sigma E_T \]
\end{enumerate}

The calibration constant \( k \) is empirically determined through a series of calibration procedures, van der Meer scans. The constant accounts for the proportionality between the aggregate transverse energy and the actual luminosity of the particle collisions.

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/luminometer_cms.png}
\caption[CMS luminometers]{%
   CMS detector with all of the luminosity measurement devices showing PLT, BCM1F, HF, DT, PCC and RAMSES \cite{meyer2021luminosity}. 
}
\label{fig:lumino_cms}
\end{figure}

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/luminometer_cms_pseudo.png}
\caption[CMS luminometers location]{%
    Cross section of the CMS detector in the r-z plane showing all Run 2 luminometers PLT, BCM1F, DTs, and HF location. The two RAMSES monitors used as a luminometer in Run 2 are located directly behind HF \cite{Sirunyan:2759951}. 
}
\label{fig:lumino_cms_pseudo}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/phaseI_upgrade_pixel_detector.png}
\caption[Pixel detector layout]{%
   Layout of the CMS Phase-1 upgraded pixel detector in longitudinal view \cite{Adam:2748381}. 
}
\label{fig:phaseI_upgrade}
\end{figure}

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/silicon_pixel_module.png}
\caption[Pixel module design]{%
   Pixel detector modules for BPIX L1 (left), BPIX L2-4 (middle), and
the FPIX detector (right) \cite{Adam:2748381}. 
}
\label{fig:pixelmodule}
\end{figure}

### 2024-03-27 tesis_Ashish/Chapter3/chapter3.tex
update 28 Mar

\begin{figure}[h]
    \centering
    \includegraphics[width=1\textwidth]{ashish_thesis/vdm_method_new.png}
    \caption[vdM Scan Schematic]{Left: Scanning the beams across each other (±3$\sigma$) in the transverse plane, horizontal (x-scan) and vertical (y-scan) directions. Right: Rate is measured as a function of beam separation \cite{karacheban2017luminosity}.}
    \label{fig:vdm_scan_method}
\end{figure}

>>>>

\begin{figure}[h]
    \centering
    \includegraphics[width=1\textwidth]{ashish_thesis/vdm_method_new.png}
    \caption[vdM scan schematic]{Left: Scanning the beams across each other (±3$\sigma$) in the transverse plane, horizontal (x-scan) and vertical (y-scan) directions. Right: Rate is measured as a function of beam separation \cite{karacheban2017luminosity}.}
    \label{fig:vdm_scan_method}
\end{figure}
